### 2012-06-20 inst/examples/Reed.Rmd
Reed exploring
Reed with uniform small noise

``` {r  echo=FALSE }
opts_knit$set(upload.fun = socialR::flickr.url)
#options(device = function(width = 5, height = 5) {
#    pdf(NULL, width = width, height = height)
#})
````

# Reed Model
 * author Carl Boettiger, <cboettig@gmail.com>
 * license: CC0

 Implements a numerical version of the SDP described in:

   Reed, W.J., 1979. Optimal Escapement Levels in Stochastic
   and Deterministic Harvesting Models. Journal of Environmental 
   Economics and Management. 6: 350-363.

Clear the workspace and load package dependencies: 
``` {r  setup, echo=FALSE}
rm(list=ls())   
require(pdgControl)
require(reshape2)
require(ggplot2)
require(data.table)
````

### Define parameters

``` {r }
gridsize <- 100   # gridsize (discretized population)
sigma_g <- 0.2    # Noise in population growth
sigma_m <- 0.     # noise in stock assessment measurement
sigma_i <- 0.     # noise in implementation of the quota
````

we'll use log normal noise functions. 
For Reed, only `z_g` will be random.
Sethi et al will add the other forms

``` {r  noise_dists}
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
z_m <- function() rlnorm(1,  0, sigma_m) # mean 1
z_i <- function() rlnorm(1,  0, sigma_i) # mean 1
````

@@

``` {r }
f <- RickerAllee
K <- 4 
xT <- 2 # final value, also allee threshold
pars <- c(1.5, K, xT) 
````

and we use a harvest-based profit function with default parameters

``` {r  profit}
profit <- profit_harvest(price=1, c0 = 0.01, c1=0) 
````

The profit_harvest function has the form \( \Pi = h - \left( c_0  + c_1 \frac{h}{x} \right) \frac{h}{x} \), conditioned on \( h > x \) and \(x > 0 \).  Note that the R code defines a function from another function using a trick known as a _closure_.  Again we could write a custom profit function as long as it can take a vector stock size `x` and a scalar harvest level `h`.  Details for provided functions can be found in the manual, i.e. `?profit_harvest`.

Now we must set up the discrete grids for stock size and havest levels (which will use same resolution as for stock), in order to calculate the SDP solution.   Here we set the gridsize to 100.

``` {r  grid}
x_grid <- seq(0, 1.5 * K, length = 100)  
h_grid <- x_grid  
````

### Calculate the transition matrix (with noise in growth only)      
We calculate the stochastic transition matrix for the probability of going from any state \(x_t \) to any other state \(x_{t+1}\) the following year, for each possible choice of harvest \( h_t \).  This provides a look-up table for the dynamic programming calculations.  Take a look at the R code and the documentation for `determine_SDP_matrix` to see what this function is actually doing.  The implementation is quite simple.

``` {r  SDP_Mat}
SDP_Mat <- determine_SDP_matrix(f, pars, x_grid, h_grid, sigma_g )
````

@@

In the Sethi case, computing the distribution over multiple sources of noise is actually quite difficult.  Simulation turns out to be more efficient than numerically integrating over each distribution.  This code parallelizes the operation over four cores, but can be scaled to an arbitrary cluster. Since we're focused on the Reed example for the moment, we can ignore this step.

``` {r  parallel, eval=FALSE}
require(snowfall) 
sfInit(parallel=TRUE, cpu=4)
sethi_SDP_Mat <- SDP_by_simulation(f, pars, x_grid, h_grid, z_g, z_m, z_i, reps=999)
sethi_opt <- find_dp_optim(sethi_SDP_Mat, x_grid, h_grid, OptTime=25, xT=0, 
                     profit, delta=0.05, reward=0)
qplot(x_grid, sethi_opt$D[,1], xlab="stock size", ylab="harvest")
````

@@

>>>>

``` {r  echo=FALSE }
opts_knit$set(upload.fun = socialR::flickr.url)
orig = getOptions("device")
options(device = function(width = 5, height = 5) {
    pdf(NULL, width = width, height = height)
})
knitcitations::cleanbib()
````

# Reed Model

 * Author [Carl Boettiger](http://carlboettiger.info), <cboettig@gmail.com>
 * License: [CC0](http://creativecommons.org/publicdomain/zero/1.0/)
 * Description:  Implements a numerical version of the SDP described in `r knitcitations::citet("10.1016/0095-0696(79)90014-7")`.

``` {r  setup, echo=FALSE, cache=FALSE}
# load required libraries
require(pdgControl)
require(reshape2)
require(ggplot2)
require(data.table)
rm(list=ls())
````

@@

``` {r }
#f <- function(x,h,p){
#	S = x - h
#	p[1] * S * (1 - S/p[2]) + S
#}
````

With parameters `r` = `r r <- 1` and `K` = `r K <- 100`.

``` {r }
#pars <- c(r, K)
````

## OR

We consider a Beverton Holt state equation governing population dynamics, \\( f(x,h) = \frac{A x}{1 + B x} \\)

``` {r }
f <- BevHolt
````

With parameters `A` = `r A <- 1.5` and `B` = `r B <- 0.05`.

``` {r }
pars <- c(1.5, 0.05)
K <- (pars[1] - 1)/pars[2]
````

We consider a profits from fishing to be a function of harvest `h` and stock size `x`,

<div> $$ \Pi(x,h) = h - \left( c_0  + c_1 \frac{h}{x} \right) \frac{h}{x}, $$ </div>

conditioned on h > x and x > 0,

``` {r  profit}
price <- 1
c0 <- 0.0
c1 <- 0
profit <- profit_harvest(price=price, c0 = c0, c1=c1) 
````

with price = `r price`, `c0` = `r c0` and `c1` = `r c1`.

``` {r }
xmin <- 0
xmax <- 1.5 * K
grid_n <- 100
````

We seek a harvest policy which maximizes the discounted profit from the fishery using a stochastic dynamic programming approach over a discrete grid of stock sizes from `r xmin` to `r xmax` on a grid of `r grid_n` points, and over an identical discrete grid of possible harvest values.

``` {r  grid}
x_grid <- seq(xmin, xmax, length = grid_n)  
h_grid <- x_grid  
````

``` {r miscpars}
delta <- 0.05
xT <- 0
OptTime <- 25
sigma_g <- .1
````

We will determine the optimal solution over a `r OptTime` time step window with boundary condition for stock at `r xT` and discounting rate of `r delta`.  The Reed model considers a stochastic growth model

<div> $$ x_{t+1} = z_g f(x_t) $$ </div>

for the random variable `z_g`, given by

``` {r noise}
z_g <- function() 1+(2*runif(1, 0,  1)-1) * sigma_g
````

No other sources of noise enter into the dynamics.

``` {r othernoise}
z_m <- function() 1
z_i <- function() 1
````

``` {r  SDP_Mat}
pdfn <- function(P, s){
  dunif(P, 1 - s, 1 + s)
}
SDP_Mat <- determine_SDP_matrix(f, pars, x_grid, h_grid, sigma_g, pdfn)
````

@@

@@

# References

``` {r biblio, echo=FALSE, results="asis"}
knitcitations::bibliography("html")
````

``` {r cleanup}
options(device=orig)
````
